---
title: "Calculating Crop Percentages and Predicting Price"
author: "Harshad B., Kuljot C. and Riya S."
output: pdf_document
---

# PART 1 : Calculating Crop Percentages

The dataset used in this example was obtained from <https://aps.dac.gov.in>. The dataset contains information of area under a crop and its production from 2010-2018 for all districts in the state of Maharashtra. The dataset contains a wide variety of crops.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
dataset <- read.csv("D:/project-alt/modded_dataset.csv")
colnames(dataset) <- c("Crop","District","Year","Season","Area","Production","Yield")
```
``` {r dataset_head}
head(dataset)
```

## Example 

In this example, we will be looking at Ahmednagar district, and the season "Rabi" for the year 2010.

```{r example}
dist_in <- "Ahmednagar"
season_in <- "Rabi"
year_in <- 2010
```   

We will subset the dataset according to our inputs.

```{r subset, include=FALSE}
data_dist_year_season <- dataset %>%
  filter(District == dist_in, Season == season_in, Year == year_in) %>%
  select(Crop, Year, Season, `Area`, `Production`, Yield) %>%
  mutate(Crop = as.factor(Crop))
crop_unique <- as.vector(unique(data_dist_year_season$Crop))
```

We will write a function to calculate the percentages of crop under the particular selections.

``` {r, include=FALSE}
crop_percent <- function(x) {
  dat <- data_dist_year_season %>%
    filter(Crop == crop_unique[x])
  prod_crop <- dat$Production
  prod_total <- sum(data_dist_year_season$Production)
  (prod_crop/prod_total)*100
}

ans <- sapply(c(1:length(crop_unique)), crop_percent)
ans_df <- data.frame(Crop = crop_unique, Percent = ans)
```

Now, the results:
```{r}
ans_df %>% 
  arrange(Percent, desc = TRUE)
```

# PART 2 : Predicting Crop Prices

The dataset used in this example was obtained from <http://agmarknet.gov.in/>. The dataset contains information of the pricing of Bajra (Pearl Millets) in Nashik District and its various markets from 2010-2018 for all dates. The dataset contains minimum, maximum and modal prices of the Bajra crop in Nashik District for the given period.

```{r setup1, include=FALSE}
library(knitr)
library(caret)
library(tidyverse)
library(data.table)
library(shiny)
library(tseries)
library(xts)
library(zoo)
library(quantmod)
library(neuralnet)

suppressWarnings(set.seed(1, sample.kind = "Rounding"))

#read data
data <- read.csv("D:/r-projects/crop_yield_prediction/ds/agricultural_price.csv")
```

Now, we will take a sneak peek at the data and while also clean the data (pre-processing) and change market names to numerical values for easier handling of data.

``` {r dataset_head1}
#taking a look at data before preprocessing
head(data)
#convert dates from int to date
data$Price.Date <- as.Date(as.character(data$Price.Date), "%d%m%Y")
class(data$Price.Date)
#remove NA dates
data$Price.Date[as.character(data$Price.Date) == 'NA'] <- NA
na_index <- which(is.na(data$Price.Date))
data_new <- data[-na_index,]
#change colnames
data_new <- data_new %>%
  rename(Sr.No = ï..Sr..No,
         District = District.Name,
         Market = Market.Name,
         MinPrice = Min.Price..Rs..Quintal.,
         MaxPrice = Max.Price..Rs..Quintal.,
         ModalPrice = Modal.Price..Rs..Quintal.,
         Date = Price.Date) %>%
  mutate(Market = as.numeric(factor(Market))) %>%
  select(Market, MinPrice, MaxPrice, ModalPrice, Date) %>%
  arrange(Date)
#take a peek at the dataframe
head(data_new)
```

The market codes are as follows:
``` {r}
markets <- data.frame(Market = unique(data$Market.Name), Factor = 1:length(unique(data$Market.Name)))
markets
```

## Example 

In this example, we will be looking at Chandvad market.

```{r example1}
target_market <- 1 #Chandvad
```   

We will subset the dataset according to our inputs.

```{r subset1, include=TRUE}
tm_data <- data_new %>%
  filter(Market == target_market)
```

We will train a neural network model with 15 hidden nodes and a threshold of 0.001.

``` {r, include=FALSE}
#extract values
Modal_data <- (tm_data$ModalPrice)
Max_data <- (tm_data$MaxPrice)
Min_data <- (tm_data$MinPrice)
#make datasets
tm_df <- cbind(Max_data,Min_data,Modal_data)
tm_actual <- data.frame(tm_df)
train_price <- (tm_actual[-(nrow(tm_actual)),])
trainingoutput_price <- as.data.frame(tm_actual[-1,3])
#Column bind the data into one variable
trainingdata_price <- as.data.frame(cbind(train_price,trainingoutput_price))
```

``` {r}
net_price <- neuralnet(tm_actual[-1, 3] ~  Max_data + Min_data, data = trainingdata_price, err.fct = "sse", hidden = 15, threshold=0.001)
```

And extract the modal pricing of the latest date available to us in the dataset.
```{r}
testdata <- as.data.frame(tm_actual[nrow(tm_actual),])
net.results <- neuralnet::compute(net_price, testdata)
Predicted_Modal_Rate<- net.results$net.result
Modal_rate<- Predicted_Modal_Rate[1,1]
```

Now, the results
```{r}
results <- data.frame(actual = testdata$Modal_data, prediction = net.results$net.result)
results
```